# Case Study1: Heart Disease Prediction Model

**Group Members:** `Manoj Paudel, Kushal Pandey, Kabin Shrestha, Rojilnem Shrestha, Sailendra Ghale`

### Importing Libraries

```{r}
library(tidyverse)
library(randomForest)
```

### Load dataset

```{r}
setwd('/Users/sandip/Documents/Github/Heart-Disease-Prediction-R/')
heart_df <- read.csv("heart.csv")
head(heart_df)
```

```{r}
glimpse(heart_df)
```

-   There are 1025 rows with 14 columns and all variable are ***int*** dtype.

```{r}
is.null(heart_df)

```

-   No missing value in dataset

    ```{r}
    is_duplicate <- duplicated(heart_df)
    is_duplicate

    sum(is_duplicate)

    # which_row_duplicate <- which(duplicated(heart_df))
    # which_row_duplicate
    ```

-   There are 723 duplicate value present in our dataset. We gonna remove those duplicates,

    ```{r}
    df <- heart_df[!is_duplicate, ]
    df

    glimpse(df)
    ```

-   Now, we have unique ***302 values***

### Is target column balance or imbalance?

```{r}
dist_target_column <- table(df$target)
dist_target_column 
```

```{r}
ggplot(df, aes(x = factor(target))) +
  geom_bar() +
  labs(title = "Distribution of Target Variable",
       x = "Target Column",
       y = "Count")
```

-   As, 0 ( No disease) have 138 value and 1 ( Disease) have 164 value, there is no significant difference in distribution so, **balance**.
