# Case Study1: Heart Disease Prediction Model

**Group Members:** `Manoj Paudel, Kushal Pandey, Kabin Shrestha, Rojil Shrestha, Sailendra Ghale`

### Objective:

To predict whether a person has a heart disease or not.

### Importing Libraries

```{r}
library(tidyverse)
library(randomForest)
library(class)
library(dplyr)
library(ggplot2)
library(GGally)
library(corrplot)
library(caTools)
library(caret)
library(ggcorrplot)

```

### Load dataset

```{r}
setwd('/Users/sandip/Documents/Github/Heart-Disease-Prediction-R/')
heart_df <- read.csv("heart.csv")
head(heart_df)

```

### Descriptions of data set

+--------------+---------------------------------------------------------------------------------+
| Features     | Descriptions                                                                    |
+==============+=================================================================================+
| **age**      | Age of the patient in years                                                     |
+--------------+---------------------------------------------------------------------------------+
| **sex**      | Gender of the patient (0 = male, 1 = female)                                    |
+--------------+---------------------------------------------------------------------------------+
| **cp**       | Chest pain type:\                                                               |
|              | 0: Typical angina\                                                              |
|              | 1: Atypical angina\                                                             |
|              | 2: Non-anginal pain\                                                            |
|              | 3: Asymptomatic                                                                 |
+--------------+---------------------------------------------------------------------------------+
| **trestbps** | Resting blood pressure in mm Hg                                                 |
+--------------+---------------------------------------------------------------------------------+
| **chol**     | Serum cholesterol in mg/dl                                                      |
+--------------+---------------------------------------------------------------------------------+
| **fbs**      | Fasting blood sugar level, categorized as above 120 mg/dl (1 = true, 0 = false) |
+--------------+---------------------------------------------------------------------------------+
| **restecg**  | Resting electrocardiographic results:\                                          |
|              | 0: Normal\                                                                      |
|              | 1: Having ST-T wave abnormality\                                                |
|              | 2: Showing probable or definite left ventricular hypertrophy                    |
+--------------+---------------------------------------------------------------------------------+
| **thalach**  | Maximum heart rate achieved during a stress test                                |
+--------------+---------------------------------------------------------------------------------+
| **exang**    | Exercise-induced angina (1 = yes, 0 = no)                                       |
+--------------+---------------------------------------------------------------------------------+
| **oldpeak**  | ST depression induced by exercise relative to rest                              |
+--------------+---------------------------------------------------------------------------------+
| **slope**    | Slope of the peak exercise ST segment:\                                         |
|              | 0: Upsloping\                                                                   |
|              | 1: Flat\                                                                        |
|              | 2: Downsloping                                                                  |
+--------------+---------------------------------------------------------------------------------+
| **ca**       | Number of major vessels (0-4) colored by fluoroscopy                            |
+--------------+---------------------------------------------------------------------------------+
| **thal**     | Thalium stress test result:\                                                    |
|              | 0: Normal\                                                                      |
|              | 1: Fixed defect\                                                                |
|              | 2: Reversible defect\                                                           |
|              | 3: Not described                                                                |
+--------------+---------------------------------------------------------------------------------+
| **target**   | Heart disease status (0 = no disease, 1 = presence of disease)                  |
+--------------+---------------------------------------------------------------------------------+
| **chol**     | Serum cholesterol in mg/dl                                                      |
+--------------+---------------------------------------------------------------------------------+

```{r}
glimpse(heart_df)
```

-   There are 1025 rows with 14 columns.

```{r}
is.null(heart_df)

```

-   No missing value in our data set.

    ```{r}
    is_duplicate <- duplicated(heart_df)
    is_duplicate

    sum(is_duplicate)

    # which_row_duplicate <- which(duplicated(heart_df))
    # which_row_duplicate
    ```

-   There are 723 duplicate value present in our data set. We gonna remove those duplicates,

    ```{r}
    df <- heart_df[!is_duplicate, ]
    # df

    glimpse(df)
    ```

-   Now, we have unique ***302 values***

------------------------------------------------------------------------

### Data Transformation

-   As, **9 columns** (`sex`, `cp`, `fbs`, `restecg`, `exang`, `slope`, `ca`, `thal`, and `target`) are **numerical** in terms of data type, but they are **categorical** features, so converting data type to factor.

```{r}
# Sex column
df$sex <- as.factor(df$sex)
# cp
df$cp <- as.factor(df$cp)
#fbs
df$fbs <- as.factor(df$fbs)
#restecg
df$restecg <- as.factor(df$restecg)
#exang
df$exang <- as.factor(df$exang)
#slope
df$slope <- as.factor(df$slope)
#ca
df$ca <- as.factor(df$ca)
#thal
df$thal <- as.factor(df$thal)
#target
df$target <- as.factor(df$target)
```

```{r}
glimpse(df)
```

```{r}
summary(df)
```

### Numerical Feature

-   **Age:** Min: 29 Max: 77 with average 55.5

-   **trestbps:** Ranging from 94 to 200 mm Hg with average 131.62 mm Hg

-   **chol:** Min 126 to max: 564 mg/dl with average 246.5 mg/dl

-   **thalach:** Ranging from 71 to 202 with average heart rate is 149.6

-   **oldpeak:** Ranging from 0 to 6.2 with average ST depression is 1.04

### Categorical Feature

-   **sex :** Male (0) 96 and female(1) 206

-   **cp:** 4 unique type of chest pain with most common 0 occuring 143 times.

-   **fbs:** Most frequent 0(low) with 257 (fasting blood sugar)

-   **restecg:** 3 unique with most common 1 with 151 times.

-   **exng:** 2 unique with most frequent 0 with 203 times.

-   **slope:** 3 unique with most frequent 2 with 141 times.

-   **ca:** 5 unique with 0 most frequent with 175 times.

-   **thal:** 4 unique with most common 2 with 165 times.

-   **target:** 0: 138 (No disease) \| 1: 164 (Presence of disease)

### Is target column balance or imbalance?

```{r}
dist_target_column <- table(df$target)
dist_target_column 
```

```{r}
ggplot(df, aes(x = factor(target))) +
  geom_bar() +
  labs(title = "Distribution of Target Variable",
       x = "Target Column",
       y = "Count")
```

-   As, 0 ( No disease) have 138 value and 1 ( Disease) have 164 value, there is no significant difference in distribution so, it is **balance**.

------------------------------------------------------------------------

### EDA

#### 1. Age vs Target column

```{r}

ggplot(df, mapping = aes(x = age, fill = target)) +
  geom_histogram() +
  facet_wrap(vars(target)) +
  labs(title = "Age vs Target", x = "Age", y = "Count", fill = "Target")

```

-   As, above plot, age between 40 to 60 have high probability of having a heart disease as compare to younger age.

#### 2. CP vs Target

```{r}
ggplot(df, mapping = aes(x=target, fill = cp)) +
  geom_bar(position = "dodge") +
  labs(title = "cp vs target", x = "Target", y = "Count", fill = "cp type")
```

-   Here, 2 (non-anginal chest pain) have highest count for a person to have a heart disease, while 3 (asymptomatic) have lowest count.

-   There is high count of 0 (typical angina chest pain) for non heart disease in comparison with those who have heart disease.

-   There is relationship between chest pain column and target column.

#### 3. Sex vs Target

```{r}
ggplot(df, mapping = aes(x = sex, fill = target)) +
  geom_bar(position = "fill") +
  labs(x = "Sex", y = "Proportion", fill = "Target", title = 'Sex vs Target') +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))
```

-   In above bar plot, male(0) have high probability of heart disease in comparison to female (1). There is relation between sex and target class.

### 4. fbs vs target

```{r}
# 0 low | 1 high

ggplot(df, mapping = aes(x = fbs, fill = target)) +
  geom_bar(position = "fill") +
  labs(x = "Fasting blood sugar", y = "Proportion", fill = "Target") +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))
```

-   There is similar distribution between (1) high blood sugar \> 120 mg/dl and (0) is relatively similar, so fbs column have no correlation with target class.

#### 5. exang vs target

```{r}
# 1 = yes, 0 = no

ggplot(df, mapping = aes(x = exang, fill = target)) +
  geom_bar(position = "fill") +
  labs(x = "Exercise Induced Angina", y = "Proportion", fill = "Target") +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))

```

-   Here, patient who did not have exercise induced angina 0 (no) show high probability of having heart disease compare to those who experience (1). There seems correlation between target class.

#### 6. slope vs target

```{r}
ggplot(df, mapping = aes(x = slope, fill = target)) +
  geom_bar(position = "fill") +
  labs(x = "slope of peak exercise(ST)", y = "Proportion", fill = "Target") +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))
```

-   Here, 2 (Down sloping) has probability of have heart disease with low probability (flat), showing potential to predict target class.

#### 7. thal vs target

```{r}
ggplot(df, mapping = aes(x = thal, fill = target)) +
  geom_bar(position = "fill") +
  labs(x = "Thalium Stress Test Result", y = "Proportion", fill = "Target") +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))
```

-   Here, 2 type (thal) have high proportion of heart disease compare to other types.

#### ca vs target

```{r}
ggplot(df, mapping = aes(x = target, fill = ca)) +
  geom_bar(position = "fill") +
  labs(x = "Target", y = "Proportion", fill = "Number of Major Vessels") +
  theme(axis.text.x = element_text(size = 12), axis.title.x = element_text(size = 12), 
        axis.title.y = element_text(size = 12), axis.text.y = element_text(size = 12))
```

-   Here, patients with 0 major vessel has highest probability of heart disease and as the major vessels increases, probability of having heart disease decreases.

**Observation:** Features ca, cp, sex, exang, slope, thal have high impact on target column.

------------------------------------------------------------------------

### Correlation Matrix

```{r}
# Only numeric value correlation
ggcorr(df, label = TRUE,label_size = 3, hjust = 1)

```

-   Above, numeric feature correlation, shows there is negative correlation with oldpeak and thalach indicate (one feature increases, other tends to decrease).

-   Correlation of all features

```{r}
# Convert factor variables back to numeric representation

# df$sex <- as.numeric(df$sex)
# df$cp <- as.numeric(df$cp)
# df$fbs <- as.numeric(df$fbs)
# df$restecg <- as.numeric(df$restecg)
# df$exang <- as.numeric(df$exang)
# df$slope <- as.numeric(df$slope)
# df$ca <- as.numeric(df$ca)
# df$thal <- as.numeric(df$thal)
# df$target <- as.numeric(df$target)
# 
# 
# cor_matrix <- cor(df, use="complete.obs")
# corrplot(cor_matrix, method="color")

```

-   We can say, ca, cp, exang, sex, slope, thal have high correlation with target class along with moderate impact by rsetecg column.

### Model Building

-   We split 70% for training set and 30% for testing set.

```{r}
set.seed(100) # random number 
train_index<-sample(nrow(df),0.70*nrow(df))
train_data <- df[train_index, ]
test_data <- df[-train_index, ]


```

### Logistic Regression

```{r}
# Train logistic regression model
logistic_model <- glm(target ~ ., data = train_data, family = "binomial")

summary(logistic_model)

```

-   **Age** shows positive coefficient, showing as age increase, the chances of having heart disease also increased.

-   **sex** with negative coefficient, 1: Female , showing being female means low chances of having heart disease in comparsion to male.

-   **cp** with positive relationship with target column.

-   **trestbps** and **chol** have negative coefficient, as there value high means lower chances of having heart disease.

-   **exang1** meaning when angina present, there is lower chances of having heart disease.

```{r}
# Evaluate models
logistic_predictions <- predict(logistic_model, newdata = test_data, type = "response")

# Convert predictions to numeric
logistic_predictions <- as.numeric(as.character(logistic_predictions))

# Convert to binary 0 and 1 of predictions
logistic_predictions <- ifelse(logistic_predictions > 0.5, 1, 0)

## Accuracy of Model
logistic_accuracy <- mean(logistic_predictions == test_data$target)
cat("Logistic Regression Accuracy:", logistic_accuracy, "\n") 
```

```{r}
confusion_matrix <- confusionMatrix(factor(logistic_predictions), factor(test_data$target))

# Print confusion matrix
print(confusion_matrix)
```

-   Using **logistic regression** we achieved accuracy with 74 %

### KNN

```{r}
## K-value calculation
sqrt(nrow(train_data))
```

-   We got 14.5258 which is even value, so we take k value as 15.

```{r}
knn_model <- knn(train = train_data[, -ncol(train_data)], test = test_data[, -ncol(test_data)], cl = train_data$target, k = 15)

# Model accurary
knn_accuracy <- mean(knn_model == test_data$target)
cat("KNN Accuracy:", knn_accuracy, "\n")
```

```{r}

confusion_matrix <- confusionMatrix(factor(knn_model), factor(test_data$target))

confusion_matrix

```

-   Using **K-Nearest** **Neighbor** we achieved accuracy with 68 % which less than **logistic regression** model.

### Random Forest

```{r}

# Training model
random_forest_model <- randomForest(target ~.,  data = train_data)

rf_predictions <- predict(random_forest_model, newdata = test_data)

# Evaluate model

rf_accuracy <- mean(rf_predictions == test_data$target)
cat("Random Forest Accuracy:", rf_accuracy, "\n")


```

```{r}
# Confusion Matrix
confusion_matrix <- confusionMatrix(rf_predictions, test_data$target)

confusion_matrix


```

-   We got accuracy of 81% using Random Forest model (highest among 3 model)

### Conclusion:

-   Based on the performance of 3 model [Random Forest, Logistic Regression, K-NN], we got accuracy of 74% with Logistic Regression and Random Forest with highest accuracy of 81% suggesting its robustness in capturing complex interactions and nonlinear relationships among features. While KNN model achieved with lowest accuracy of 68 % which is way less in comparison of 2 other model.
